A Triple Origin for Twin Blue Stragglers in Close Binaries

### 2019-01-16 BSS7782.tex
updated the paper

\def\apgt{\ {\raise-.5ex\hbox{$\buildrel>\over\sim$}}\ }
\def\aplt{\ {\raise-.5ex\hbox{$\buildrel<\over\sim$}}\ }

@@

We propose a formation mechanism for twin blue stragglers in compact
binaries that involves mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
this scenario to the observed double blue straggler system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin blue
stragglers: (1) For the outer tertiary orbit, the initial orbital
period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$
1100 days, assuming initial masses for the inner binary components of
$m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After
Roche-lobe overflow, the outer star turns into a white dwarf (WD) of
0.43 to 0.54\,\MSun. There is a correlation between the mass of this
WD and the outer orbital period in the sense that more massive white
dwarfs will be on wider orbits.  (3) The rotational axes of both the
BSs in the inner binary will be aligned with each other and the
orbital plane of the outer tertiary WD. (4) The BSs in the inner
binary will have roughly the same masses, independent of their initial
masses.  This follows from the more copious accretion of mass onto the
initially lower mass star.  This star should therefore be enriched
more effectively by the accreted material.  As a result, one of the
blue stragglers will appear to be more enriched by He, C and O
if the donor started to overflow its Roche lobe at the red
giant branch, or with s-process elements if the donor was on the
asymptotic giant branch.  (5) Relative to old dense clusters with high
velocity dispersions, twin BSs in close binaries formed from the
proposed mechanism should be more frequent in the Galactic field and
younger open clusters with ages $\lesssim$ 4-6 Gyr, since then the
donor will have a radiative envelope. (6) the orbit of the binary BS
will be close (typically $\aplt 0.3$\,au) and close to circular ($e
\aplt 0.2$).

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,2014ApJ...782...49S}.  Two
primary channels for BS formation have been proposed: mass transfer
from an evolved donor on to a main-sequence star in a binary star
system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving main-sequence stars possibly
mediated via binaries
\citep[e.g.][]{hills75,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H,2018arXiv181100058P}.
Other possible, albeit related, formation mechanism includes mergers
of close main-sequence binaries (MS-MS) \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.

@@

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner binary composed to two main-sequence
stars.  In section~\ref{sect:dyn}, we constrain the range of initial
(i.e., pre-mass transfer) orbital parameters for a hypothetical outer
tertiary companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process triple
system with orbital parameters that, according to our expectations,
for promising progenitors for the twin BS 7782.  The calculations are
performed using the Astrophysical Multipurpose Software Environment
(\texttt{AMUSE} for short) \cite{AMUSE} with a combination of stellar
evolution, hydrodynamical and gravitational simulations.  with these
calculations we further constrain the possible range of initial
parameters that naturally lead to twin BS with orbital parameters
similar to the 7782 system.  We summarize and discuss the implications
of our results for compact double BS binaries and, more generally,
mass transfer in stellar triples in Section~\ref{sect:discussion}.

@@

In our scenario we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we now
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\cite{2010yCat..73890925T,moe18}.  This initial configuration for our
assumed formation scenario for Binary 7782, described below, is
depicted in Figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it leaves the asumptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamicall stable, for which we adopt eq.\,90 in
\cite{2001MNRAS.321..398M}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk.  Using conservation of
angular momentum, we equate the specific angular momentum of the
accreted mass at the inner Lagrangian point of the (outer) donor star
to the final specific angular momentum of the accretion stream at the
circularization radius about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the center of mass corresponding to the tertiary defined by the
Roche lobe is given by Equation 2 in \citep{1983ApJ...268..368E}.
Combining Equation 2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
Equation~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopted
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin blue stragger 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green colored curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertikcal dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (brown)
    shows an estimate of the final orbit separation of the outer star,
    and therefore of the final orbit of the white dwarf around the
    inner twin blue straggler.  For core masses $\apgt 0.5$\,\MSun\,
    the final orbital separataion, after mass transfer, is smaller
    than the initial orbit.  Here we adopted an initial inner binary
    mass of (1.0+0.9)\,\MSun\, and a final twin-blue straggler mass of
    (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after the phase of
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a'$) and after mass transfer
($a$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a \over a'} = \left( {m'_3 m'_{\rm in} \over m_3 m_{\rm in}} \right)^{-2}
  \left( {m'_3 + m'_{\rm in} \over m_3 + m_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides an
estimate of the orbital separation of the white dwarf around the twin
blue straggler 7782. Tidal effects during the episode of mass tranfer
have probably circularized the orbit, althoug some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small eccentricity of $e \ll 0.1$.

@@

We perform simulations of a triple star system for which the outer
star overfills its Roche lobe while the inner binary remains
detached. The calculations start by evolving the three stars to the
same age, which is selected such that the outer-most star fills its
Roche lobe.  First order constraints for the initial conditions are
derived in the previous \S. In the following two sections we describe
how we setup these simulations and then discuss the results. The
calculations are performed using the

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components, and
between $m_{\rm 3} = 1.2$ and $m_{\rm 3} = 1.4$ M$_{\rm \odot}$ for
the tertiary star.  We evolve the tertiary star using the MESA
stellar-evolution code \cite{2011ApJS..192....3P} to a radius of about
1\,au, at which point we assume it to overfill it's Roche lobe.  By
this time the star has lost 0.13\,\MSun\, via a stellar wind.

The stellar-evolution model, including the structure, temperature and
composition profiles, are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills its
Roche lobe and trnsfers mass to an inner binary.  After generating the hydrodynamical
representation of the donor star we replace the stellar core by a
point mass to prevent the majority of the resolution to be confined in
the star's central regions.  In a following step we relax the star
using the hydrodynamics solver. This relaxation process is realized in
100 steps during which we reduce the velocity dispersion of individual
SPH particles to a glasseas structure \citep[see, for example, \S\,3.3
  on page 40 in][]{1994astro.ph.10043W}. During this procedure the
gaseous envelope of the star tends to expand by about 20\%.  To
detemrine the radius of the evolving star we calculate Lagrangian
radii and use the distance to the stellar center which contains 90\%
of it's mass. From this 90\% mass-radius relation we obtain the stellar radius,
and match it with the Roche-lobe of the outer orbit.

With these parameters the orbital separation of the outer binary
becomes $\sim 3.86$\,au, and we adopt the outer orbit to be circular.
We choose a relative initial inclination between the inner and the
outer orbits of $i = 0^\circ$, $i = 9^\circ$, $i = 90^\circ$ and $i =
180^\circ$.  The first two choices are motivated by observed low-mass stellar
triples composed of low-mass stars which have been observed to favor
co-planar configurations \citep{2018ApJ...854...44M}. To test the effect of highly
inclined and retrogade orbits we include the $i = 90^\circ$ and $i =
180^\circ$ cases.  In Fig.\,\ref{fig:topview_at_t0} we present a top
view of the triple system in our simulations at the onset of Roche-lobe overflow.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t0002.pdf}
\caption{Top view of the simulated triple system in which the outer
  star over-fills its Roche-lobe (represented by 80000\,sph particles
  and a core particle of 0.4\,\MSun (blue bullet). The two companion
  stars are represented as black bullets (to the right).  The inner
  binary is represented by the yellow and the red bullets for the
  1.1\,\MSun\, primary and 0.9\,\MSun\, secondary stars in a circular
  orbit of 0.1\,au. The 1.4\,\MSun\, giant star is presented to the
  right in a circular orbit of $\sim 1.47$\,au in the plan of the
  inner binary.
\label{fig:topview_at_t0}}
\end{figure}

Roche-lobe overflow in triples is modeled using a coupled integrator
to follow the complex hydrodynamics of mass transfer from the
Roche-lobe filling outer star to the inner binary, while keeping track
of the gravitational dynamics of the stars.  The equations of motion
of the inner binary are solved using the symplectic direct N-body
integrator \texttt{Huayno} \citep{2012NewA...17..711P}. The
hydrodynamics are performed with the smoothed-particles hydrodynamics
code \texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete mass and angular momentum from
the gas liberated by the outer star.  This is realized using
sink-particles that co-move with the mass points in the gravity
code. While the inner two stars accrete mass, they also accrete the
corresponding amount of angular momentum from the gas (see chapter 5
in \cite{AMUSE}).  The N-body integrator correctly accounts for this.  For the radius of the sink particles we adopt $10
R_\odot$ for both stars.

The N-body code as well as the hydrodynamics solver operate using
their own internal time-steps. The coupling between the two codes is
realized using the \texttt{Bridge} method in the AMUSE framework
\citep[see Sect.\.4.3.1 in][]{2013CoPhC.183..456P}.  This coupled
integrator is based on the splitting of the Hamiltonian, much in the
same way as is done with two different gravity solvers by
\cite{2007PASJ...59.1095F}. With the adopted scheme, the
hydrodynamical solver is affected by the gravitational potential of
its own particles, as well as the gravitational potential of the inner
binary. The hydrodynamics, in particular the interaction with the gas,
in turn affects the orbits of the two inner stars. With
\texttt{Bridge} we realize a second order coupling between the
gravitational dynamics and the hydrodynamics.  The interval at which
the gravity and hydrodynamics interact via \texttt{Bridge} depends on
the parameters of the system we study, but typically we achieve
converged solutions when this time step is about a fraction of 1/100
of the inner binary orbital period.

\section{Results} \label{results}

In fig.\,\ref{fig:topview_at_t1000day} we show the same triple after
1000\,days of evolution.
%NL2:  We are labeling the figures differently in the text.  You label with ``fig.\,\", whereas I use ``Figure~''.  It doesn't matter to me which we use, just thought I'd highlight this so we come back to it right before submitting.

\cite{2018arXiv181208175M}

\begin{figure}[ht!]
  \includegraphics[width=0.49\columnwidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.pdf}
~  \includegraphics[width=0.51\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t3360.pdf}
\caption{Same as Fig.\,\ref{fig:topview_at_t0} but at an age of $t
  \simeq 1091\,days$ after the start of the simulation. Left panel
  shows the equipotential surface of the triple overplotted with the
  gas distribution, the right panel shows just the gas and the stars
  as bullets.
\label{fig:topview_at_t1000day}}
\end{figure}

In fig.\,\ref{fig:mass_vs_semimajor_axis}

@@

In this Letter, we have proposed a formation scenario for double BS
equal-mass compact binaries, as observed for Binary 7782 in the old
open cluster NGC 188.  The proposed scenario involves mass transfer
from an evolved outer tertiary companion, which is accreted by the
inner binary via a circumbinary disk.  Our scenario makes several
predictions for the observed properties of a hypothetical outer triple
companion, now a WD.  These are:

>>>>

\def\apgt{\ {\raise-.5ex\hbox{$\buildrel>\over\sim$}}\ }
\def\aplt{\ {\raise-.5ex\hbox{$\buildrel<\over\sim$}}\ }
%%\def\figure{\ {figure\,}}

@@

We propose a formation mechanism for twin blue stragglers in compact
binaries that involve mass transfer from an evolved outer tertiary
companion on to the inner binary via a circumbinary disk.  We apply
this scenario to the observed double blue straggler system Binary 7782
in the old open cluster NGC 188, and show that its observed properties
are naturally reproduced within the context of the proposed model.
Based on this model, we predict the following properties for twin blue
stragglers: (1) For the outer tertiary orbit, the initial orbital
period should lie between 220 days $\lesssim$ P$_{\rm out}$ $\lesssim$
1100 days, assuming initial masses for the inner binary components of
$m_{\rm 1} = 1.1$ M$_{\odot}$ and $m_{\rm 2} =$ 0.9 M$_{\odot}$ and an
outer tertiary mass of $m_{\rm 3} = 1.4$ M$_{\odot}$.  After
Roche-lobe overflow, the outer star turns into a white dwarf (WD) of
0.43 to 0.54\,\MSun. There is a correlation between the mass of this
WD and the outer orbital period in the sense that more massive white
dwarfs will be on wider orbits.  (3) The rotational axes of both the
BSs in the inner binary will be aligned with each other and the
orbital plane of the outer tertiary WD. (4) The BSs in the inner
binary will have roughly the same masses, independent of their initial
masses.  This follows from the more copious accretion of mass onto the
initially lower mass star.  This star should, therefore, be enriched
more effectively by the accreted material.  As a result, one of the
blue stragglers will appear to be more enriched by He, C and O
if the donor started to overflow its Roche lobe at the red
giant branch, or with s-process elements, if the donor was on the
asymptotic giant branch.  (5) Relative to old dense clusters with high-velocity dispersions, twin BSs in close binaries formed from the
proposed mechanism should be more frequent in the Galactic field and
younger open clusters with ages $\lesssim$ 4-6 Gyr, since then the
donor will have a radiative envelope. (6) the orbit of the binary BS
will be close (typically $\aplt 0.3$\,au) and close to circular ($e
\aplt 0.2$).

@@

Blue straggler (BS) stars are brighter and bluer than the
main-sequence turn-off in a cluster colour-magnitude diagram
\citep[e.g.][]{1953AJ.....58...61S,2014ApJ...782...49S}.  Two primary
channels for BS formation have been proposed: mass transfer from an
evolved donor on to a main-sequence star in a binary star system
\citep[e.g.][]{1964MNRAS.128..147M,1997A&A...328..143P,2009Natur.457..288K,2011MNRAS.410.2370L,2011Natur.478..356G},
and direct stellar collisions involving main-sequence stars possibly
mediated via binaries
\citep[e.g.][]{1975AJ.....80..809H,1997A&A...328..130P,2007ApJ...661..210L,2013MNRAS.428..897L,2013MNRAS.429.1221H,2018arXiv181100058P}.
Other possible, albeit related, formation mechanism includes mergers
of close main-sequence binaries (MS-MS) \cite{2018arXiv181100058P},
and mergers of the inner binaries of hierarchical triple star systems
induced by Lidov-Kozai oscillations coupled with tidal damping
\citep[e.g.][]{2009ApJ...697.1048P}.

@@

We propose a formation channel for Binary 7782, and compact double BS
binaries in general, which involves mass transfer from an outer
tertiary companion on to an inner binary composed of two main-sequence
stars.  In section~\ref{sect:dyn}, we constrain the range of initial
(i.e., pre-mass transfer) orbital parameters for a hypothetical outer
tertiary companion, using a combination of dynamical and stellar
evolution-based constraints.  In Section~\ref{sims} we present the
numerical simulations used to study the mass transfer process triple
system with orbital parameters that, according to our expectations,
for promising progenitors for the twin BS 7782.  The calculations are
performed using the Astrophysical Multipurpose Software Environment
(\texttt{AMUSE} for short) \cite{AMUSE} with a combination of stellar
evolution, hydrodynamical and gravitational simulations.  with these
calculations we further constrain the possible range of initial
parameters that naturally lead to twin BS with orbital parameters
similar to the 7782 system.  We summarize and discuss the implications
of our results for compact double BS binaries and, more generally,
mass transfer in stellar triples in Section~\ref{sect:discussion}.

@@

In our scenario, we start with a binary star with component masses
$m_{\rm 1}$ and $m_{\rm 2}$ that is orbited tertiary of mass
$m_3$. The inner and outer binary orbital semi-major axes are denoted
a$_{\rm in}$ and a$_{\rm out}$, respectively.  For clarity we now
assume both orbits, the inner as well as the outer, to have negligible
eccentricity and low inclination.  These assumptions are also
supported by the population of observed low-mass triples
\cite{2010yCat..73890925T,moe18}.  This initial configuration for our
assumed formation scenario for Binary 7782, described below, is
depicted in figure~\ref{fig:fig1}.

@@

According to our scenario $m_3 > m_1 > m_2$ and the outer orbit is
sufficiently small that the tertiary star is filling its Roche lobe
and transfers mass to the inner binary before it leaves the asymptotic
giant branch. We constrain the inner orbit by requiring the triple
system to be dynamically stable, for which we adopt eq.\,1 in
\cite{1999ASIC..522..385M,}.  While transferring mass, the accretion
stream gathers around the inner binary at the circularization radius
a$_{\rm c}$, and forms a circumbinary disk
\citep{2002apa..book.....F}.  Using conservation of angular momentum,
we equate the specific angular momentum of the accreted mass at the
inner Lagrangian point of the (outer) donor star to the final specific
angular momentum of the accretion stream at the circularization radius
about the inner binary, this results in
\begin{equation}
\label{eqn:specangmom1}
v_{\rm orb,3}(a_{\rm out} - R_{\rm L}) = v_{\rm orb,c}a_{\rm c},
\end{equation}
where R$_{\rm L}$ is the radius of the Roche lobe of the outer
tertiary companion, $a_{\rm c}$ is the semi-major axis of the orbit
about the inner binary corresponding to the circularization radius and
v$_{\rm orb,c}$ is the orbital velocity at $a_{\rm c}$.  The distance
from the centre of mass corresponding to the tertiary defined by the
Roche lobe is given by Equation 2 in \citep{1983ApJ...268..368E}.
Combining Equation 2 in \citet{1983ApJ...268..368E} (with mass ratio q
$=$ m$_{\rm 3}$/(m$_{\rm 1} +$m$_{\rm 2}$)) with
Equation~\ref{eqn:specangmom1}, we solve for the circularization
radius as a function of a$_{\rm out}$ and the assumed stellar masses:
\begin{equation}
\label{eqn:ac}
a_{\rm c} = a_{\rm out}(1 - R_{\rm L}).
\end{equation}
In order for a circumbinary disk to form around the inner binary, we
require that a$_{\rm in} <$ a$_{\rm c}$.

@@

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_M14MSun_outerorbit.pdf}
  \caption{Giant radius as a function of the mass of the core of the
    Roche-lobe filling outer star (dark blue curve).  Here we adopted
    a donor mass of 1.4\,\MSun, but for an 1.2\,\MSun\, donor curve is
    quite similar (see dotted dark-blue curve).  The red squares in
    the curve show the parameters for which we performed more detailed
    gravitational-hydrodynamical simulations (see \S\,\ref{sims}).
    The horizontal dashed line shows the orbital separation of the
    observed twin blue straggler 7782.  The initial triple in which it
    possibly formed must at least have been dynamically stable. The
    minimal orbital separation for the inner binary for which the
    triple is stable is given by the lower green coloured curve.
    Donors which are smaller than about 100\,\RSun\, (dark blue curve
    indicated with a$_{\rm stable}$) result in a dynamically unstable
    triple. The minimal core mass associated with a stable triple is
    then indicated by the left-most vertical dotted line.  The
    orbital separation at which the donor star overfills its Roche
    lobe is indicated with the light-blue curve. The top curve (brown)
    shows an estimate of the final orbit separation of the outer star,
    and therefore of the final orbit of the white dwarf around the
    inner twin blue straggler.  For core masses $\apgt 0.5$\,\MSun\,
    the final orbital separation, after mass transfer, is smaller
    than the initial orbit.  Here we adopted an initial inner binary
    mass of (1.0+0.9)\,\MSun\, and a final twin-blue straggler mass of
    (1.4+1.4)\,\MSun.
\label{fig:tertiarymass_vs_size}}
\end{figure}

@@

With the adopted parameters, we can also estimate the final orbital
period of the left-over core from the tertiary star after the phase of
mass transfer.  The change in orbital separation due to
non-conservative mass transfer can be expressed in terms of the mass
of the outer star before and after mass transfer $m_3$ and $m'_3$,
respectively, the total mass in the inner binary before ($m_{\rm in}$)
and after accretion ($m'_{\rm in}$) and the amount of angular momentum
lost per unit mass $\eta \simeq 3$. Adopting the relation between the
orbital separation before mass transfer ($a$) and after mass transfer
($a'$) from \cite{1995A&A...296..691P}
\begin{equation}
  {a' \over a} = \left( {m_3 m_{\rm in} \over m'_3 m'_{\rm in}} \right)^{-2}
  \left( {m_3 + m_{\rm in} \over m'_3 + m'_{\rm in}} \right)^{2\eta + 1},
\end{equation}
we arrive at the top brown curve in
figure\,\ref{fig:tertiarymass_vs_size}. This curve provides an
estimate of the orbital separation of the white dwarf around the twin
blue straggler 7782. Tidal effects during the episode of mass tranfer
have probably circularized the orbit, althoug some slight eccentricity
due to turbulent motion in the outer layers of the donor star may have
induces a small $e \ll 0.1$ eccentricity.

Having limited parameter space for the formation of the twin blue
straggler 7782, we continue by performing a series of simulations to
investigate the accretion and changes to the inner orbits of triple
systems in this range of parameters.

@@

We perform simulations of a triple star system for which the outer
star overfills its Roche lobe while the inner binary remains
detached. The calculations start by evolving the three stars to the
same age, which is selected such that the outer-most star fills its
Roche lobe.  First order constraints for the initial conditions are
derived in the previous \S. In the following two sections we describe
how we set up these simulations and then discuss the results. The
calculations are performed using the Astrophysical Multipurpose
Software Environment using a combination of stellar evolution,
gravitational dynamics and hydrodynamics.

@@

We adopt initial masses of $m_{\rm 1} = 1.1$ M$_{\rm \odot}$ and
$m_{\rm 2} = 0.9$ M$_{\rm \odot}$ for the inner binary components, and
between $m_{\rm 3} = 1.2$ and $m_{\rm 3} = 1.4$ M$_{\rm \odot}$ for
the tertiary star.  We evolve the tertiary star using the MESA
stellar-evolution code \cite{2011ApJS..192....3P} to a radius of about
100\,\RSun\, and 150\,\RSun, at which point we assume it to overfill
it's Roche lobe (see red squared in
figure~\ref{fig:tertiarymass_vs_size}).  We perform calculations for an
inner orbital separation of $a_{\rm in} = 0.10$\,au, $a_{\rm in} =
0.15$\,au and $a_{\rm in} = 0.20$\,au.  In total we performed 12
calculations at a resolution of 40k SPH particles 12 at 80k.

The stellar-evolution model, including the structure, temperature and
composition profiles are turned into a smoothed-particles
representation using the module {\tt StellarModelInSPH} in AMUSE (see
chapter 4 in \cite{AMUSE}).  We follow the same procedure as described
in \cite{2014MNRAS.438.1909D} for simulating the future of the triple
system $\chi$ Tau (HD 97131) in which the outer-most star overfills
its Roche lobe and transfers mass to an inner binary.  After
generating the hydrodynamical representation of the donor star we
replace the stellar core by a point mass to prevent the majority of
the resolution to be confined in the star's central regions.  In a
following step we relax the star using the hydrodynamics solver. This
relaxation process is realized in 100 steps during which we reduce the
velocity dispersion of individual SPH particles to a glasses
structure \citep[see, for example, \S\,3.3 on page 40
  in][]{1994astro.ph.10043W}. During this procedure, the gaseous
envelope of the star tends to expand by about 20\%.  To determine the
radius of the evolving star we calculate Lagrangian radii and use the
distance to the stellar center which contains 90\% of its mass. From
this 90\% mass-radius relation we obtain the stellar radius and match
it with the Roche-lobe of the outer orbit.

With these parameters the orbital separation of the outer binary
becomes $\sim 250$\,\RSun for the 100\,\RSun donor star and about
430\,\RSun\, for the more evolved donor star.  We adopt the outer
orbit to be circular and in the plane of the inner binary.
%% We
%% performed several additional calculations for inclined orbits (choosing
%% a relative initial inclination $i = 9^\circ$, $i = 45^\circ$, $i =
%% 90^\circ$ and $i = 180^\circ$, but we wi
The low inclination is motivated by observed stellar triples composed
of low-mass stars to favour co-planar configurations
\citep{2018ApJ...854...44M}.
%% To test the effect of highly inclined
%% and retrogade orbits we include the $i = 90^\circ$ and $i = 180^\circ$
%% cases.
In figure~\ref{fig:topview_at_t0} we present a top view of the initial
conditions for one of these calculations.

\begin{figure}[ht!]
  \includegraphics[width=\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t0002.pdf}
\caption{Top view of the simulated triple system in which the
  100\,\RSun\, outer star of 1.4\,\MSun\, that over-fills its
  Roche-lobe.  the star is represented by 80000\,SPH particles and a
  core particle of $\sim 0.4$\,\MSun (black bullet). The two companion
  stars are represented as black bullets (to the right).  The inner
  binary is represented by the yellow and the red bullets for the
  1.1\,\MSun\, primary and 0.9\,\MSun\, secondary stars in a circular
  orbit of 0.1\,au. The 1.4\,\MSun\, giant star is presented to the
  right in a circular orbit of $\sim 250$\,\RSun\, in the plan of the
  inner binary.
\label{fig:topview_at_t0}}
\end{figure}

Roche-lobe overflow in triples is modelled using a coupled integrator
to follow the complex hydrodynamics of mass transfer from the
Roche-lobe filling outer star to the inner binary, while keeping track
of the gravitational dynamics of the stars.  The equations of motion
of the inner binary are solved using the symplectic direct N-body
integrator \texttt{Huayno} \citep{2012NewA...17..711P}. The
hydrodynamics are performed with the smoothed-particles hydrodynamics
code \texttt{Gadget2} \citep{2000ascl.soft03001S}, using an adiabatic
equation of state.  The two inner binary stars are treated as point
masses, but we allow them to accrete mass and angular momentum from
the gas liberated by the outer star.  This is realized using spherical
sink-particles that co-move with the mass points in the gravity
code. While the inner two stars accrete mass, they also accrete the
corresponding amount of angular momentum from the gas (see chapter 5
in \cite{AMUSE}).  The N-body integrator correctly accounts for this.
For the radius of the sink particles, we adopt $2 R_\odot$ for both
stars.

The N-body code, as well as the hydrodynamics solver, operate using
their own internal time-steps. The coupling between the two codes is
realized using the \texttt{Bridge} method in the AMUSE framework
\citep[see Sect.\.4.3.1 in][]{2013CoPhC.183..456P}.  This coupled
integrator is based on the splitting of the Hamiltonian, much in the
same way as is done with two different gravity solvers by
\cite{2007PASJ...59.1095F}. With the adopted scheme, the
hydrodynamical solver is affected by the gravitational potential of
its own particles, as well as the gravitational potential of the inner
binary. The hydrodynamics, in particular, the interaction with the gas,
in turn, affects the orbits of the two inner stars. With
\texttt{Bridge} we realize a second order coupling between the
gravitational dynamics and the hydrodynamics.  The interval at which
the gravity and hydrodynamics interact via \texttt{Bridge} depends on
the parameters of the system we study, but typically we achieve
converged solutions when this time step is about a fraction of 1/100
of the inner binary orbital period.

\section{Results of the hydrodynamical simulations} \label{results}

to test the hypotheses that (1) the secondary in the inner binary
accretes more effectively than the primary star and to measure the
change to the inner orbit due to the Roche-lobe overflow of the outer
star, we perform a series of calculations in which we take the self
gravity and the hydrodynamical effects of the triple into account.
The results of these simulations are presented in
figure~\ref{fig:topview_at_t1000day} and
figure~\ref{fig:mass_vs_semimajor_axis}.  The first figure
(figure~\ref{fig:topview_at_t1000day} shows the top view of the same
initial realization for which we presented the initial conditions in
figure~\ref{fig:topview_at_t0} but now at an age of 1091 days after
the onset of mass transfer. To guide the eye we add, to the left panel,
the equipotential surfaces in the orbital plane.

It is apparent that the mass transfer in the adopted triples leads to
a rather untidy evolution, in particular, because much of the donor
mass is lost through the second Lagrangian point to the right side of
the donor star in figure~\ref{fig:topview_at_t1000day}. Much mass is
also lost through the third Lagrangian point (to the left of the inner
binary). One remaining question is how much mass is eventually ejected
altogether from the triple system and is therefore not accreted to any
of the two inner stars. This value is hard to estimate from the
simulations, but an accretion efficiency of $\apgt 0.6$ is necessary to
make the scenario feasible. Over the time scale for which we performed
the calculations, this efficiency is reached, but it is not clear who
the system responds at later stages.

\begin{figure}[ht!]
  \includegraphics[width=0.49\columnwidth]{fig_BBSS_Lgas_M14Msun_A010au_t3360.pdf}
~  \includegraphics[width=0.51\columnwidth]{fig_BBSS_gas_M14Msun_A010au_t3360.pdf}
\caption{Same as figure~\ref{fig:topview_at_t0} but at an age of $t
  \simeq 1091\,days$ after the start of the simulation. Left panel
  shows the equipotential surface of the triple overplotted with the
  gas distribution, the right panel shows just the gas and the stars
  as bullets.
\label{fig:topview_at_t1000day}}
\end{figure}

The evolution of the inner orbit presented for several simulations in
figure~\ref{fig:mass_vs_semimajor_axis}, is complicated.  This is
caused by the complex transport of mass, energy and angular momentum
through the accretion stream and throughout the system.  It is
therefore hard to quantify distinct trends in the evolution of the
triple system. In simulated of the response of an inner binary on
accretion from a circumbinary disk, \cite{2018arXiv181208175M}
conclude that the complexity of angular momentum transport between
outer star, accretion stream to the individual inner stars, is
complicated and without clear trends. For most of our calculations we
agree with this statement, but in
figure~\,\ref{fig:mass_vs_semimajor_axis} we nevertheless present the
results of 6 of our calculations, three for and 1.2\,\MSun\, donor and
three for a 1.4\,\MSun\, donor star. The various coloured curves give
the resulting evolution of the inner orbit as a function of the total
mass in the inner binary. While the inner two stars accrete the orbit
in the case with a 1.2\,\MSun\, donor tends to shrink. These systems
are expected to result in a contact binary, that eventually may merge
to form a single blue straggler. The required evolution in order to
explain the observed twin blue straggler 7782 are indicated with the
three black curves, and the simulated path clearly deviates from
these. We, therefore, argue that a 1.2\,\MSun\, donor has difficulty
explaining the observed orbital separation of $\sim 0.13$\,au in BSS
7782.

@@

We propose a formation scenario for twin blue stragglers of equal-mass
in tight binaries, as observed for Binary 7782 in the old open cluster
NGC 188.  The proposed scenario involves mass transfer from an evolved
outer tertiary companion, which is accreted by the inner binary via a
circumbinary disk.  Our scenario makes several predictions for the
observed properties of a hypothetical outer triple companion, now a
WD.  These are:
